---
title: "Structural Bioinformatics (pt. 1)"
author: "Jack Reddan"
output: github_document
---

# Introduction to the RCSB Protein Data Bank (PDB)  
  
## PDB Statistics  
  
```{r}
pdb_stats <- read.csv("pdb_statistics.csv")

pdb_stats
```
  
### [Q01]: What percentage of structures in the PDB are solved by X-Ray and Electron Microscopy?  
  
```{r}
# x_ray_percent <- round(((sum(pdb_stats$X.ray) / sum(pdb_stats$Total)) * 100), 2)
# paste0("X-Ray: ", x_ray_percent, "%", sep = "")
# 
# EM_percent <- round(((sum(pdb_stats$EM) / sum(pdb_stats$Total)) * 100), 2)
# paste0("XElectron Microscopy: ", EM_percent, "%", sep = "")

round((colSums(pdb_stats[,2:(ncol(pdb_stats)-1)])/sum(pdb_stats$Total)*100), 2)
```
  
### [Q02]: What percentage of structures in the PDB are protein?  
  
```{r}
protein_percent <-round(
  (sum(pdb_stats$Total[grep("Protein", pdb_stats$Molecular.Type)])/sum(pdb_stats$Total)*100), 
  2)
paste0("Proteins: ", protein_percent, "%", sep = "")
```
  
### [Q03]: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?  
  
2343 structures.  
  
## The PDB Format  

### [Q04]: Water molecules normally have 3 atoms. Why do we see just one atom per water molecule in this structure?  
  
The hydrogen atom is extremely small and is not depicted because the oxygen atom
sufficiently identifies the location and spatial occupation of the water molecule.
  
### [Q05]: There is a conserved water molecule in the binding site. Can you identify this water molecule? What residue number does this water molecule have (see note below)?  
  
HOH 308:O
