---
title: "Lab 06"
author: "Jack Reddan (A59010543)"
date: "10/15/2021"
output: pdf_document
#output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Question 1:
### Write a function grade() to determine an overall grade from a vector of student homework assignment scores dropping the lowest single score.
This function will provide the grade for a student given a vector of their scores,
and will provide the average of all scores while dropping the lowest score.

```{r}
grade <- function(scores) {
  scores <- as.numeric(scores)
  scores[is.na(scores)] <- 0
  scores <- scores[-which.min(scores)]
  return(mean(scores))
}
```

When tested on these student cases, the *grade* function should result in the 
average score of student1 without the **90**, student2 without the **NA** and student3
without one **NA** and the rest of the NAs represented as a 0.

```{r}
# Example input vectors to start with
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)

#Test 'grade' function
c("Student 1" = grade(student1), 
  "Student 2" = grade(student2), 
  "Student 3" = grade(student3))
```

```{r}
#gradebook <- read.csv(file = "student_homework.csv", row.names = 1)
gradebook <- read.csv(file = "https://tinyurl.com/gradeinput", row.names = 1)
```

## Question 2:
### Who is the top scoring student overall in the gradebook?
Apply the grade function across all students in the gradebook dataframe. Then,
identify the student with the maximum calculated grade average.

```{r}
adjusted_averages <- apply(gradebook, 1, grade)
adjusted_averages[which.max(adjusted_averages)]
```

## Question 3:
### Which homework was toughest on students?
Generate a copy of the gradebook dataframe, and convert all *NA* values to 0.

```{r}
gradebook_mask <- gradebook
gradebook_mask[is.na(gradebook_mask)] <- 0
```

Apply the mean function across the homeworks in the gradebook dataframe. Then,
identify the homework with the minimum calculated grade average.

```{r}
hw_averages <- apply(gradebook_mask, 2, mean)
hw_averages[which.min(hw_averages)]
```

## Question 4:
### Which homework was most predictive of overall score?
Using the *cor* function,
calculate the correlation between each student's score for a homework assignment 
and the average score of each student for that homework assignment.
Then,
identify the homework with the strongest correlation.

```{r}
homework_correlations <- cor(gradebook_mask, adjusted_averages)
homework_correlations[which.max(homework_correlations),1]
```

