---
title: "Mini Project"
author: 'Jack Reddan (PID: A59010543)'
date: "10/27/2021"
#output: pdf_document
output: github_document
---
## Organizing the data  

First I read in the data from 'WisconsinCancer.csv" with the first column as the row names.  

```{r}
fna_data <- "WisconsinCancer.csv"
wisconsin_df <- read.csv(fna_data, row.names = 1)
```

Then I look at the data to asses the structure of the data.  

```{r}
str(wisconsin_df)
```
  
The first row contains the diagnoses which are the outputs we are trying to
derive from the data. This will be removed to generate the naive dataframe.
Additionally, these results are saved as the vector 'diagnosis'.

```{r}
naive_wisconsin_df <- wisconsin_df[,-1]
diagnosis <- factor(wisconsin_df[,1])
```

## Explore the data

>[Q1]: How many observations are in the dataset?  

```{r}
nrow(wisconsin_df)
```
  
There are 569 rows/observations.  
  
>[Q2]: How many of the observations have a malignant diagnoses?  
  
```{r}
sum(diagnosis == "M")
```
  
212 of the observations have a "M" or malignant diagnoses.  
  
>[Q3]: 