---
title: "Introduction to Genome Informatics"
author: "Jack Reddan (PID: A59010543)"
output: pdf_document
---

# Questions  
  
#### Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.  
  
#### Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?  
  
# Answers
   
## Upload the Data  
  
```{r}
res <- read.table("expression_genotype_results")
res$geno <- as.factor(res$geno)
```
  
## Explore the Data  
  
```{r}
geno_distro <- boxplot(exp ~ geno,
        data = res,
        xlab = "Genotype",
        ylab = "ORMDL3 Expression",
        notch = TRUE,
        col = c("#FFB000", "#648FFF", "#DC267F"),
        main = "Gene expression levels for SNPs")
```
  
Median values for genotypes `r geno_distro$names` are `r geno_distro$stats[3,]`, respectively. 
The population sizes are `r geno_distro$n`.  
  
Based off the boxplot, it appears that ORMDL3 gene expression is highest when homozygous for the A SNP (A/A), and lowest when homozygous for the G SNP (G/G). 
Additionally, it appears to be dose dependent since gene expression levels are intermediate for the heterozygote (A/G). 
Therefore, the SNP does seem to affect ORMDL3 gene expression, where the more copies of the A SNP present leads to higher ORMDL3 expression.

