---
title: "Halloween Mini-Project"
author: 'Jack Reddan (PID: A59010543)'
date: "10/29/2021"
#output: html_document
output: pdf_document
---

# Library loading  
```{r}
library(skimr)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(corrplot)
library(plotly)
```
  
# Importing candy data  
  
```{r}
candy_file <- "candy-data.csv"
candy <- read.csv(candy_file, row.names = 1)
head(candy)
```

##### [Q1]: 
How many different candy types are in this dataset?  
  
```{r}
nrow(candy)
```
  
##### [Q2]: 
How many fruity candy types are in the dataset?
  
```{r}
sum(candy$fruity)
```

##### [Q3]: 
What is your favorite candy in the dataset and what is it’s `winpercent`
value?  
  
```{r}
candy["100 Grand",]$winpercent
```

##### [Q4]: 
What is the `winpercent` value of "Kit Kat"?  
  
```{r}
candy["Kit Kat",]$winpercent
```
  
##### [Q5]: 
What is the `winpercent` value for "Tootsie Roll Snack Bars"  
  
```{r}
candy["Tootsie Roll Snack Bars",]$winpercent

```
  
```{r}
#skim(candy)
```

##### [Q6]: 
Is there any variable/column that looks to be on a different scale to the
majority of the other columns in the dataset?  
  
Yes, the 'winpercent' column, the mean is two orders of magnitude larger than the
others.  
  
##### [Q7]: 
What do you think a zero and one represent for the `candy$chocolate` 
column?  

1 and 0 represent TRUE and FALSE for whether the category could be used to
describe the candy of interest.  
  
##### [Q8]: 
Plot a histogram of `winpercent` values.  
  
```{r}
hist(x = candy$winpercent)
```
  
##### [Q9]: 
Is the distribution of winpercent values symmetrical?  
  
  
```{r}
hist(x = candy$winpercent)
abline(v = mean(candy$winpercent), col = 'red')
abline(v = median(candy$winpercent), col = 'blue')
```
  
No, the data has a higher mean than median, indicating more density at the 
higher end of the distribution. Therefore, the distribution of the data is not 
symmetrical.  
  
##### [Q10]: 
Is the center of the distribution above or below 50%?  
  
*See plot above* The median/center of the distribution is below 50%.
  
##### [Q11]: 
On average is chocolate candy higher or lower ranked than fruit candy?  
  
```{r}
fruity_logi <- as.logical(candy$fruity)
choco_logi <- as.logical(candy$chocolate)

fruity_win <- candy$winpercent[fruity_logi]
names(fruity_win) <- rownames(candy)[fruity_logi]

choco_win <- candy$winpercent[choco_logi]
names(choco_win) <- candy$winpercent[choco_logi]

mean(choco_win)
mean(fruity_win)
```
  
On average, chocolate candy is ranked higher, 60.92%, compared to fruity candy
(44.12%).  
  
##### [Q12]: 
Is this difference statistically significant?  
  
```{r}
choco_vs_furity_stat <- t.test(choco_win, fruity_win)
choco_vs_furity_stat$p.value
```
  
Yes, the difference is significant with a p-value of 2.87x10^-8.  
  
# Overall Candy Rankings  
  
##### [Q13]: 
What are the five least liked candy types in this set?  
  
```{r}
candy %>% arrange(winpercent) %>% head(5)
```

##### [Q14]: 
What are the top 5 all time favorite candy types out of this set?  
  
```{r}
candy %>% arrange(desc(winpercent)) %>% head(5)
```
  
##### [Q15]: 
Make a first barplot of candy ranking based on winpercent values.  
  
```{r}
ggplot(data = candy) +
  aes(x = winpercent, y = rownames(candy)) +
  geom_col()
```
  
  
##### [Q16]: 
This is quite ugly, use the `reorder()` function to get the bars sorted 
by `winpercent`?  
  
```{r}
ggplot(data = candy) +
  aes(x = winpercent, y = reorder(rownames(candy), winpercent)) +
  geom_col()
```

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```
  
```{r}
ggplot(data = candy) +
  aes(x = winpercent, y = reorder(rownames(candy), winpercent)) +
  geom_col(fill = my_cols)
```
  
##### [Q17]: 
What is the worst ranked chocolate candy?  

Boston Baked Beans.  
  
##### [Q18]: 
What is the best ranked fruity candy?  
  
Starbursts.   
  
# Taking a look at pricepercent  
  
```{r}
ggplot(candy) + 
  aes(winpercent, pricepercent, label = rownames(candy)) +
  geom_point(col = my_cols) +
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```
  
##### [Q19]: 
Which candy type is the highest ranked in terms of winpercent for the 
least money - i.e. offers the most bang for your buck?  
  
Reese's Miniatures.  
  
##### [Q20]: 
What are the top 5 most expensive candy types in the dataset and of 
these which is the least popular?  
  
```{r}
candy %>% 
  arrange(desc(pricepercent)) %>% 
  head(5)
```
   
```{r}
candy %>% 
  arrange(desc(pricepercent)) %>% 
  head(5) %>% 
  arrange(winpercent) %>% 
  head(1)
```

   
Nik L Nip is the most expensive of the least favorite candy options.  
  
##### [Q21]: 
Make a barplot again with geom_col() this time using pricepercent.

```{r}
ggplot(data = candy) +
  aes(x = pricepercent, y = reorder(rownames(candy), pricepercent)) +
  geom_col()
```
  
```{r}
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), xend = 0), 
               col="gray40") +
    geom_point()
```

# Exploring the Correlation Structure  
  
```{r}
candy_cor <- cor(candy)
corrplot(candy_cor)
```

##### [Q22]: 
Examining this plot what two variables are anti-correlated (i.e. have 
minus values)?  
  
Chocolate and fruity.  
  
##### [Q23]: 
Similarly, what two variables are most positively correlated?  
  
Winpercent and chocolate.  
  
# Principal Component Analysis

```{r}
candy_pca <- prcomp(candy, scale = TRUE)
summary(candy_pca)
```
  
```{r}
plot(candy_pca$x[,1:2],
     xlab = "PC 1",
     ylab = "PC 2")
```
  
```{r}
plot(candy_pca$x[,1:2],
     col = my_cols,
     pch = 16,
     xlab = "PC 1",
     ylab = "PC 2")
```
  
```{r}
candy_df <- cbind(candy, candy_pca$x[,1:3])
```
  
```{r}
p <- ggplot(candy_df) + 
  aes(x=PC1, y=PC2, 
      size=winpercent/100,
      text=rownames(candy_df),
      label=rownames(candy_df)) + 
  geom_point(col=my_cols) +
  labs(x = "PC 1", y = "PC 2")

p
```
  
```{r}
p <- p + 
  geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") + 
  labs(title="Halloween Candy PCA Space", 
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")

p
```

<!--``{r}-->
<!--ggplotly(p)-->
<!--```-->
  
```{r}
op <- par(mar=c(8,4,2,2))
barplot(candy_pca$rotation[,1], las=2, ylab="PC1 Contribution")
par(op)
```
  
##### [Q24]: 
What original variables are picked up strongly by PC1 in the positive 
direction? Do these make sense to you?  
  
Fruity, hard and pluribus.
Yes, these variables are found together often, and can easily differentiate 
between types of candy.

# Session Information
```{r}
sessionInfo()
```

