---
title: "Halloween Mini-Project"
author: 'Jack Reddan (PID: A59010543)'
date: "10/29/2021"
output: github_document
#output: pdf_document
---

# Library loading  
```{r}
library(skimr)
library(dplyr)
```



# Importing candy data  
  
```{r}
candy_file <- "candy-data.csv"
candy <- read.csv(candy_file, row.names = 1)
head(candy)
```

>[Q1]: How many different candy types are in this dataset?  
  
```{r}
nrow(candy)
```
  
>[Q2]: How many fruity candy types are in the dataset?
  
```{r}
sum(candy$fruity)
```

>[Q3]: What is your favorite candy in the dataset and what is it’s `winpercent`
value?  
  
```{r}
candy["100 Grand",]$winpercent
```

>[Q4]: Waht is the `winpercent` value of "Kit Kat"?  
  
```{r}
candy["Kit Kat",]$winpercent
```
  
>[Q5] What is the `winpercent` value for "Tootsie Roll Snack Bars"  
  
```{r}
candy["Tootsie Roll Snack Bars",]$winpercent

```
  
```{r}
skim(candy)
```

>[Q6]: Is there any variable/column that looks to be on a different scale to the
majority of the other columns in the dataset?  
  
Yes, the 'winpercent' column, the mean is two orders of magnitude larger than the
others.  
  
>[Q7]: What do you think a zero and one represent for the `candy$chocolate` 
column?  

1 and 0 represent TRUE and FALSE for whether the category could be used to
describe the candy of interest.  
  
>[Q8]: Plot a histogram of `winpercent` values.  
  
```{r}
hist(x = candy$winpercent)
```
  
>[Q9]: Is the distribution of winpercent values symmetrical?  
  
```{r}
hist(x = candy$winpercent)
abline(v = mean(candy$winpercent), col = 'red')
abline(v = median(candy$winpercent), col = 'blue')
```
  
No, the data has a higher density below the mean, since the median is shifted 
below the mean. Therefore, the distribution of the data is not symmetrical.  
  
>[Q10]: Is the center of the distribution above or below 50%?  
  
*See plot above* The median/center of the distribution is below 50%.
  
>[Q11]: On average is chocolate candy higher or lower ranked than fruit candy?  
  
```{r}
fruity_logi <- as.logical(candy$fruity)
choco_logi <- as.logical(candy$chocolate)

fruity_win <- candy$winpercent[fruity_logi]
names(fruity_win) <- rownames(candy)[fruity_logi]

choco_win <- candy$winpercent[choco_logi]
names(choco_win) <- candy$winpercent[choco_logi]

mean(choco_win)
mean(fruity_win)
```

>[Q12]: Is this difference statistically significant?  
  
```{r}
choco_vs_furity_stat <- t.test(choco_win, fruity_win)
choco_vs_furity_stat$p.value
```
  
  
# Overall Candy Rankings  
  
>[Q13]: What are the five least liked candy types in this set?  
  
```{r}
candy %>% arrange(winpercent) %>% head(5)
```

