---
title: "Vaccination Rate Mini Project"
author: "Jack Reddan (PID: A59010543)"
output: github_document
---

# Getting Started  
  
```{r}
library(dplyr)
library(lubridate)
#library(zipcodeR)
library(ggplot2)
```
  
```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
```
  
```{r}
head(vax)
```
  
#### [Q01]: What column details the total number of people fully vaccinated?  
  
`persons_fully_vaccinated`  
  
#### [Q02]: What column details the Zip code tabulation area? 
  
`zip_code_tabulation_area`  
  
#### [Q03]: What is the earliest date in this dataset?  
  
```{r}
vax %>% 
  arrange(as_of_date) %>%
  head(1)[1]
```
  
#### [Q04]: What is the latest date in this dataset?  
  
```{r}
vax %>%
  arrange(desc(as_of_date)) %>%
  head(1)[1]
```
  
```{r}
skimr::skim(vax)
```
  
#### [Q05]: How many numeric columns are in this dataset?  
  
9  
  
#### [Q06]: Note that there are “missing values” in the dataset. How many `NA` values there in the `persons_fully_vaccinated` column?  

8355  
  
#### [Q07]: What percent of `persons_fully_vaccinated` values are missing (to 2 significant figures)?  
  
```{r}
round((1 - 0.899)*100, 2)
```
  
#### [Q08]: Why might this data be missing?  
  
  
  
# Working with Dates  
  
```{r}
today()
```
  
```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```
  
```{r}
today() - vax$as_of_date[1]
```
  
```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```
  
#### [Q09]: How many days have passed since the last update of the dataset?  
  
```{r}
last_update <- vax %>%
  arrange(desc(as_of_date)) %>%
  head(1)[1]

today() - last_update
```
  
#### [Q10]: How many unique dates are in the dataset (i.e. how many different dates are detailed)  
  
```{r}
length(unique(vax$as_of_date))
```
  
# Working with ZIP Codes  
  
```{r}
geocode_zip("92037")
```  
  
```{r}
zip_distan
```

# Focus on the San Diego Area  
  
```{r}
sd <- vax[vax$county == "San Diego",]
```  
  
With `dplyr`:  
  
```{r}
sd <- vax %>%
  filter(county == "San Diego")
```  
  
```{r}
sd.10 <- vax %>%
  filter(county == "San Diego") %>%
  filter(age5_plus_population > 10000)
```  
  
#### [Q11]: How many distinct zip codes are listed for San Diego County?  
  
```{r}
length(unique(sd$zip_code_tabulation_area))
```
  
#### [Q12]: What San Diego County Zip code area has the largest 12 + Population in this dataset?  
  
```{r}
sd %>%
  arrange(desc(age12_plus_population)) %>%
  head(1)[2]
```
  
#### [Q13]: What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09”?  
  
```{r}
sd.yest <- sd %>%
  filter(as_of_date == "2021-11-23") %>%
  filter(!is.na(percent_of_population_fully_vaccinated))

paste(round(mean(sd.yest$percent_of_population_fully_vaccinated)*100, 2), "%", sep = "")
```
  
#### [Q14]: Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”?  

```{r}
hist(sd.yest$percent_of_population_fully_vaccinated,
     xlab = "Percent of Population Fully Vaccinated on 2021-11-23",
     ylab = "Frequency",
     main = "Histogram of Vaccination Rates Across San Diego County")
```
  
# Focus on UCSD/La Jolla  
  
```{r}
ucsd <- sd %>%
  filter(zip_code_tabulation_area == "92037")

ucsd$age5_plus_population[1]
```  
  
#### [Q15]: Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:  
  
```{r}
ggplot(data = ucsd) +
  aes(x = as_of_date, 
      y = percent_of_population_with_1_plus_dose) + 
  geom_point() +
  geom_line(group = 1) +
  ylim(c(0,1)) +
  labs(x = "Date",
       y = "Percent Vaccinated")
```




  
 