---
title: "Vaccination Rate Mini Project"
author: "Jack Reddan (PID: A59010543)"
#output: github_document
output:
  pdf_document:
    latex_engine: xelatex
---

# Getting Started  
  
```{r}
library(dplyr)
library(lubridate)
#library(zipcodeR)
library(ggplot2)
```
  
```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
```
  
```{r}
head(vax)
```
  
#### [Q01]: What column details the total number of people fully vaccinated?  
  
`persons_fully_vaccinated`  
  
#### [Q02]: What column details the Zip code tabulation area? 
  
`zip_code_tabulation_area`  
  
#### [Q03]: What is the earliest date in this dataset?  
  
```{r}
vax %>% 
  arrange(as_of_date) %>%
  head(1)[1]
```
  
#### [Q04]: What is the latest date in this dataset?  
  
```{r}
vax %>%
  arrange(desc(as_of_date)) %>%
  head(1)[1]
```
  
```{r}
skimr::skim(vax)
```
  
#### [Q05]: How many numeric columns are in this dataset?  
  
9  
  
#### [Q06]: Note that there are “missing values” in the dataset. How many `NA` values there in the `persons_fully_vaccinated` column?  

8355  
  
#### [Q07]: What percent of `persons_fully_vaccinated` values are missing (to 2 significant figures)?  
  
```{r}
round((1 - 0.899)*100, 2)
```
  
#### [Q08]: Why might this data be missing?  
  
  
  
# Working with Dates  
  
```{r}
today()
```
  
```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```
  
```{r}
today() - vax$as_of_date[1]
```
  
```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```
  
#### [Q09]: How many days have passed since the last update of the dataset?  
  
```{r}
last_update <- vax %>%
  arrange(desc(as_of_date)) %>%
  head(1)[1]

today() - last_update
```
  
#### [Q10]: How many unique dates are in the dataset (i.e. how many different dates are detailed)  
  
```{r}
length(unique(vax$as_of_date))
```
  
# Working with ZIP Codes  
  
<!-->```{r}<-->
<!-->geocode_zip("92037")<-->
<!-->```<-->  
  
<!-->```{r}<-->
<!-->zip_distan<-->
<!-->```<-->

# Focus on the San Diego Area  
  
```{r}
sd <- vax[vax$county == "San Diego",]
```  
  
With `dplyr`:  
  
```{r}
sd <- vax %>%
  filter(county == "San Diego")
```  
  
```{r}
sd.10 <- vax %>%
  filter(county == "San Diego") %>%
  filter(age5_plus_population > 10000)
```  
  
#### [Q11]: How many distinct zip codes are listed for San Diego County?  
  
```{r}
length(unique(sd$zip_code_tabulation_area))
```
  
#### [Q12]: What San Diego County Zip code area has the largest 12 + Population in this dataset?  
  
```{r}
sd %>%
  arrange(desc(age12_plus_population)) %>%
  head(1)[2]
```
  
#### [Q13]: What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09”?  
  
```{r}
sd.yest <- sd %>%
  filter(as_of_date == "2021-11-23") %>%
  filter(!is.na(percent_of_population_fully_vaccinated))

paste(round(mean(sd.yest$percent_of_population_fully_vaccinated)*100, 2), "%", sep = "")
```
  
#### [Q14]: Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”?  

```{r}
hist(sd.yest$percent_of_population_fully_vaccinated,
     xlab = "Percent of Population Fully Vaccinated on 2021-11-23",
     ylab = "Frequency",
     main = "Histogram of Vaccination Rates Across San Diego County")
```
  
# Focus on UCSD/La Jolla  
  
```{r}
ucsd <- sd %>%
  filter(zip_code_tabulation_area == "92037")

ucsd$age5_plus_population[1]
```  
  
#### [Q15]: Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:  
  
```{r}
ggplot(data = ucsd) +
  aes(x = as_of_date, 
      y = percent_of_population_fully_vaccinated) + 
  geom_point() +
  geom_line(group = 1) +
  ylim(c(0,1)) +
  labs(x = "Date",
       y = "Percent Vaccinated",
       title = "Vaccination Rate of UCSD/La Jolla Zipcode")
```
  
# Compare to Similar Sized Areas  
  
```{r}
vax.lj_pop <- vax %>%
  filter(age5_plus_population >= ucsd$age5_plus_population) %>%
  filter(as_of_date == "2021-11-23")
```
  
#### [Q16]: Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-23”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?
  
The mean is `r paste(round(mean(vax.lj_pop$percent_of_population_fully_vaccinated)*100, 2), "%", sep = "")`.  
  
```{r}
ggplot(data = ucsd) +
  aes(x = as_of_date, 
      y = percent_of_population_fully_vaccinated) + 
  geom_point() +
  geom_line(group = 1) +
  geom_hline(yintercept = mean(vax.lj_pop$percent_of_population_fully_vaccinated),
            linetype = 2,
            col = "red") +
  ylim(c(0,1)) +
  labs(x = "Date",
       y = "Percent Vaccinated",
       title = "Vaccination Rate of UCSD/La Jolla Zipcode")
```
  
#### [Q17]: What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-23”?  
  
```{r}
summary(vax.lj_pop$percent_of_population_fully_vaccinated)
```
  
#### [Q18]: Using ggplot generate a histogram of this data.  
  
```{r}
ggplot(data = vax.lj_pop) + 
  aes(x = percent_of_population_fully_vaccinated) +
  geom_bar()
```
  
#### [Q20]: Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?  
  
```{r}
avg_vax_rate <- mean(vax.lj_pop$percent_of_population_fully_vaccinated)

zc_92109 <- vax %>%
  filter(as_of_date == "2021-11-23") %>%
  filter(zip_code_tabulation_area == 92109)

zc_92040 <- vax %>%
  filter(as_of_date == "2021-11-23") %>%
  filter(zip_code_tabulation_area == 92040)

```
  
```{r}
sd.lj_pop <- sd %>%
  filter(age5_plus_population >= ucsd$age5_plus_population)

length(unique(sd.lj_pop$zip_code_tabulation_area))
```
  
```{r}
ggplot(data = sd.lj_pop) +
  aes(x = as_of_date,
      y = percent_of_population_fully_vaccinated,
      group = zip_code_tabulation_area) +
  geom_line()
```
  
```{r}
vax.lj_pop_all <- vax %>%
  filter(age5_plus_population >= ucsd$age5_plus_population)

length(unique(vax.lj_pop_all$zip_code_tabulation_area))
```
  
```{r}
mean.lj_pop <- mean(vax.lj_pop$percent_of_population_fully_vaccinated)

ggplot(data = vax.lj_pop_all) +
  aes(x = as_of_date,
      y = percent_of_population_fully_vaccinated,
      group = zip_code_tabulation_area) +
  geom_line(alpha = 0.2,
            color = "blue") +
  geom_hline(yintercept = mean.lj_pop,
             color = "red",
             linetype = 2) +
  labs(x = "Date",
       y = "Percent Vaccinated",
       title = "Vaccination Rates Across California",
       subtitle = "Only areas with a population above or equalt othat of La Jolla")
```
  





  
 