---
title: "K-Means Problem"
author: "Jack Reddan"
date: "10/21/2021"
output: github_document
#output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Try K-Means clustering

Generate fake data and explore how the method works.

## Generate example data
```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))
hist(tmp)
```

## Generate multidimensional example data
```{r}
x <- cbind(x = tmp, y = rev(tmp))

plot(x)
```

## Use the kmeans() function to explore the fake data
Use it while specifying 2 expected clusters and iterating 20 times.
```{r}
clusters <- kmeans(x, centers = 2, nstart = 20)

clusters
```

### [Q] How many points are in each cluster?
There are 30 points in each cluster.
```{r}
clusters$size
```

### [Q] What component of your results object dteails:
- Cluster size
```{r}
clusters$size
```

- Cluster assignment
```{r}
clusters$cluster
```

- Cluster center
```{r}
clusters$centers
```

### Plot x colored by the kmeans cluster centers as blue points

Load ggplot2
```{r}
library(ggplot2)
```

Convert matrices to be used in ggplot to data frames.
```{r}
df <- data.frame(x)
centroids <- data.frame(clusters$centers)
```

Plot the original data colored by kmenas clusters and add blue centroids. 
IBM's colorblind palette is used.
```{r}
ggplot(data = df) +
  aes(x = x, y = y, color = factor(clusters$cluster)) +
  geom_point() +
  scale_color_manual(values = c("#785EF0", "#FE6100"), name = "Cluster") +
  geom_point(data = centroids, aes(x = x, y = y), color = "#648FFF", shape = 8)
```
