---
title: "Lab 08"
author: "Jack Reddan"
date: "10/22/2021"
output: github_document
---

## Import the Data

```{r}
url <- "https://tinyurl.com/UK-foods"
uk_foods <- read.csv(url, row.names = 1)
```

## Explore Imported Data
>Question 1: How many rows and columns are in your new data frame named x? What R functions could you use to answer this questions?  

```{r}
dim(uk_foods)
```

## Checking the Data

```{r}
head(uk_foods)
```

>Question 2: Which approach to solving the ‘row-names problem’ mentioned above do you prefer and why? Is one approach more robust than another under certain circumstances?  

I prefer to read in the row-names when I read in the csv data (in *read.csv*).
This is less dangerous since it does not mutate our data to in the body of the 
code, and instead restricts it to data initiation.

# Looking for differences

```{r}
barplot(as.matrix(uk_foods), beside = T, col = rainbow(nrow(uk_foods)))
```
>Question 3: Changing what optional argument in the above barplot() function results in the following plot?  

```{r, echo=FALSE}
barplot(as.matrix(uk_foods), beside = F, col = rainbow(nrow(uk_foods)))
```  
Changing the `beside` argument from `TRUE` to `FALSE`  

>Question 5: Generating all pairwise plots may help somewhat. Can you make sense of the following code and resulting figure? What does it mean if a given point lies on the diagonal for a given plot?  

```{r}
pairs(uk_foods, col=rainbow(10), pch=16)
```  

This code plots all pairwise comparisons of row values across the column groups.
If a given point lies on the diagonal, this means that the value of the
corresponding row is the same in both columns which are plotted. A point above
the diagonal would be higher in the column plotted on the y-axis than in the column plotted on the x-axis. A point below the diagonal would be higher in the 
column plotted on the x-axis than in the column plotted on the y-axis.  

>Question 6: What is the main differences between N. Ireland and the other countries of the UK in terms of this data-set?  

>>Without PCA  

Identify the log2 fold change and order as decreasing
```{r}
n_ireland_v_england <- log(uk_foods$N.Ireland, base = 2)/log(uk_foods$England, base = 2)
n_ireland_v_wales <- log(uk_foods$N.Ireland, base = 2)/log(uk_foods$Wales, base = 2)
n_ireland_v_scotland <- log(uk_foods$N.Ireland, base = 2)/log(uk_foods$Scotland, base = 2)

names(n_ireland_v_england) <- row.names(uk_foods)
names(n_ireland_v_wales) <- row.names(uk_foods)
names(n_ireland_v_scotland) <- row.names(uk_foods)

order(n_ireland_v_england, decreasing = T)
order(n_ireland_v_wales, decreasing = T)
order(n_ireland_v_scotland, decreasing = T)
```


## PCA

### Conduct PCA with the UKFoods data set
```{r}
uk_foods_pca <- prcomp(t(uk_foods))
summary(uk_foods_pca)
```

### Inspect PCAs 
```{r}
plot(uk_foods_pca)
```

### Plot the data using PCs
Use the two PCs which explain the most variance to plot the UKFoods data set.
```{r}
plot(uk_foods_pca$x[,1:2])
text(uk_foods_pca$x[,1:2], colnames(uk_foods))
```
```{r}
uk_foods_pca$rotation
```

```{r}
op <- par(mar=c(10,3,0.35,0))
barplot(uk_foods_pca$rotation[,1], las = 2)
par(op)
```

#PCA of RNA-Seq Data

## Import the Data
```{r}
url2 <- "https://tinyurl.com/expression-CSV"
rna_data <- read.csv(url2, row.names=1)
head(rna_data)
```

>Question 10: How many genes and samples are in this data set?  

```{r}
nrow(rna_data)
```

## PCA for RNA-Seq Data
```{r}
rna_pca <- prcomp(t(rna_data), scale = TRUE)
summary(rna_pca)
```

Scree plot of the PCs
```{r}
plot(rna_pca)
```

Plot the data using PC1 and PC2
```{r}
plot(rna_pca$x[,1], rna_pca$x[,2], xlab = "PC1", ylab = "PC2")
```

Check if the gorupings make sense
```{r}
rna_pca_variance <- rna_pca$sdev^2
rna_pca_varience_percent <- round(rna_pca_variance/sum(rna_pca_variance)*100, 1)

color_vector <- colnames(rna_data)
color_vector[grep("wt", color_vector)] <- "red"
color_vector[grep("ko", color_vector)] <- "blue"

plot(rna_pca$x[,1], rna_pca$x[,2], col=color_vector, pch=16,
     xlab=paste0("PC1 (", rna_pca_varience_percent[1], "%)"),
     ylab=paste0("PC2 (", rna_pca_varience_percent[2], "%)"))

text(rna_pca$x[,1], rna_pca$x[,2], labels = colnames(rna_data), pos=c(rep(4,5), rep(2,5)))
```
