---
title: "Lab 05"
author: "Jack Reddan (A59010543)"
date: "10/14/2021"
#output: html_document
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab 05: Data Visualization

## Install packages
```{r}
# install.packages("ggplot2")
# install.packages("gapminder")
# install.packages("patchwork")
```


## Library loading
```{r, message=FALSE}
library(ggplot2)
library(gapminder)
library(dplyr)
#library(plotly)
library(patchwork)
```

## Data exploration
Examine the *cars* dataset with the "head" function.

```{r}
head(cars)
```

## Plotting the data
Graph stopping distance (ft) as a function of speed (mph) from the 'cars' 
dataset using ggplot2.

```{r}
ggplot(data = cars) +
  aes(x = speed, y = dist) +
  geom_point() +
  geom_smooth(formula = y ~ x, 
              method = "lm") +
  labs(title = "Stopping Distance of Old Cars",
       x = "Speed (MPH)",
       y = "Stopping Distance (ft)")
```

Compare the ggplot2 graph to the graph generated from the plotting which comes
from base R.

```{r}
# Base graphics translation
plot(x = cars$speed, 
     y = cars$dist, 
     pch = 16,
     main = "Stopping Distance of Old Cars\n[base graphics]",
     xlab = "Speed (MPH)",
     ylab = "Stopping Distance (ft)")
abline(lm(cars$dist ~ cars$speed), 
       col = "blue",
       lwd = 2)
```

## Using the gene expression dataset
Reading in the gene expression tab-delimited data table as a dataframe and 
checking the table structure.

```{r}
# Loading in the genes data
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```

Exploring the data frame dimensions.

```{r}
# How many genes are in the dataframe
nrow(genes)

# What and how many types of data were collected on each gene.
colnames(genes)
ncol(genes)

# Summary of State values for each gene
table(genes$State)

# Calculate the percentage of genes which were labeled "up" for their state
round(table(genes$State)["up"]/nrow(genes)*100, 2)
```

Plotting the gene states for both treatment conditions and coloring each data 
point by its state value.

```{r}
p <- ggplot(data = genes) +
  aes(x = Condition1, y = Condition2, 
      col = State) +
  geom_point()
p
```

Adding a unique color scheme to the plot.

```{r}
p <- p +
  scale_color_manual(values = c("red", "gray", "blue"))
p  
```

Adding a title and axis labels to the plot.

```{r}
p <- p +
  labs(title = "Gene Expression Changes Upon Drug Treatment",
       x = "Control (no drug)",
       y = "Drug Treatment")
p
```

## Using the gapminder dataset
Filter the gapminder dataframe for data associated with the year 2007.

```{r}
# gapminder dataset
gapminder_2007 <- gapminder %>% filter(year==2007)
```

Plot the gapminder data for life expectancy (lifeExp) across available years.

```{r}
# Use the geom_violin and geom_jitter layers to accurately view the distribution 
# of data points
p <- ggplot(data = gapminder) +
  aes(x = year, y = lifeExp) +
  geom_violin(aes(group = year), draw_quantiles = c(0.5)) +
  geom_jitter(aes(col = continent), alpha = 0.4, width = 0.3)
p
```

<!-- Turn the plot interactive with the plotly library -->

<!-- ```{r} -->
<!-- ggplotly(p) -->
<!-- ``` -->

Plot the 2007-subset of gapminder data for life expectancy (lifeExp) across 
the countrys' associated GDP per capita.

```{r}
p <- ggplot(data = gapminder_2007) +
  aes(x = gdpPercap, y = lifeExp, col = continent, size = pop) +
  geom_point(alpha=0.4)
p
```

Using a continuous variable for color.

```{r}
# Gross example with population as color
q <- ggplot(data = gapminder_2007) +
  aes(x = gdpPercap, y = lifeExp, col = pop) +
  geom_point()
q
```

Scaling the size of the data points to better represent the differences in 
population sizes.

```{r}
# Adjusting point size
p <- ggplot(data = gapminder_2007) +
  aes(x = gdpPercap, y = lifeExp, size = pop) +
  geom_point(alpha=0.4)
p

p <- p + scale_size_area()
p
```

Filter the gapminder dataframe for data associated with the year 1957.

```{r}
# Exploring 1957 gapfinder data
gapminder_1957 <- gapminder %>% filter(year==1957)
```

Plot the 1957-subset of gapminder data for life expectancy (lifeExp) across 
the countrys' associated GDP per capita.

```{r}
p <- ggplot(data = gapminder_1957) +
  aes(x = gdpPercap, y = lifeExp, col = continent, size = pop) +
  geom_point(alpha=0.7) +
  scale_size_area(max_size = 15)
p
```

Filter the gapminder dataframe for data associated with the years 1957 and 2007.

```{r}
gapminder_1957_2007 <- gapminder %>% filter(year==1957 | year == 2007)
```

Plot the gapminder_1957_2007 data for life expectancy (lifeExp) across 
the countrys' associated GDP per capita and facet wrap by the year.

```{r}
p <- ggplot(data = gapminder_1957_2007) +
  aes(x = gdpPercap, y = lifeExp, col = continent, size = pop) +
  geom_point(alpha=0.7) +
  scale_size_area(max_size = 15) +
  facet_wrap(~year)
p
```

## Boxplots in ggplot
Filter the gapminder data for values from the year 2007 and select the top 5 
countries with the highest population

```{r}
# Top 5 population in the gapminder data for the year 2007
gapminder_2007_top5 <- gapminder %>% 
  filter(year == 2007) %>%
  arrange(desc(pop)) %>%
  top_n(5, pop)

# Check the tibble
gapminder_2007_top5
```

Plot the population for each of the top five countries with the highest populations,
and color by the country's continent.

```{r}
ggplot(data = gapminder_2007_top5) +
  aes(x = country, y = pop, fill = continent) +
  geom_col()
```

Change the color mapping to the life expectancy of the country.

```{r}
ggplot(data = gapminder_2007_top5) +
  aes(x = country, y = pop, fill = lifeExp) +
  geom_col()
```

Change the order of the countries from highest to lowest and color by country.

```{r}
ggplot(data = gapminder_2007_top5) +
  aes(x = reorder(country, -pop), y = pop, fill = country) +
  geom_col()
```

## Using the the USArrests dataset
Check the structure of the dataframe.

```{r}
head(USArrests)
```

Plot the murder arrests (per 100,000) for each state in the US, and flip the 
orientation of the plot 90 degrees.

```{r}
ggplot(data = USArrests) +
  aes(x = reorder(rownames(USArrests), Murder), y = Murder) +
  coord_flip() +
  geom_col()
```

Change the boxplot to a collection of points and line segments for easier
interpretation.

```{r}
ggplot(data = USArrests) +
  aes(x = reorder(rownames(USArrests), Murder), y = Murder) +
  geom_point() +
  geom_segment(x = rownames(USArrests),
               xend = rownames(USArrests),
               y = 0,
               yend = USArrests$Murder) +
  coord_flip()
```

## Combining plots
Assign example plots to variables.

```{r}
p1 <- ggplot(mtcars) + geom_point(aes(mpg, disp))
p2 <- ggplot(mtcars) + geom_boxplot(aes(gear, disp, group = gear))
p3 <- ggplot(mtcars) + geom_smooth(aes(disp, qsec))
p4 <- ggplot(mtcars) + geom_bar(aes(carb))
```

Combine the example plots with the *patchwork* library.

```{r}
(p1 | p2 | p3) /
      p4
```


## Session information

```{r}
sessionInfo()
```